---
title: "Extractions"
author: "Sarah Tonnar"
date: "07/07/2023"
output: html_document
---

#### Converting lion data frames to shapefiles
```{r}
# Lion 1:10
coordinates(lion10) <- c("Easting", "Northing")
proj4string(lion10) <- projection

# Lion 1:15
coordinates(lion15) <- c("Easting", "Northing")
proj4string(lion15) <- projection

# Lion 1:20
coordinates(lion20) <- c("Easting", "Northing")
proj4string(lion20) <- projection
```

#### Converting hyena data frames to shapefiles
```{r}
# Hyena 1:10
coordinates(hyena10) <- c("Easting", "Northing")
proj4string(hyena10) <- projection

# Hyena 1:15
coordinates(hyena15) <- c("Easting", "Northing")
proj4string(hyena15) <- projection

# Hyena 1:20
coordinates(hyena20) <- c("Easting", "Northing")
proj4string(hyena20) <- projection
```

***
#### Extracting the distance to the nearest waterhole

##### Creating a buffer
```{r}
# Creating a 50 metre buffer around waterholes
water_buffer <- gBuffer(waterholes, byid=TRUE, width = 50)
# Setting CRS
proj4string(water_buffer) <- projection
```

##### Lion to waterhole
```{r}
# Creating a matrix of distances between observation points and the waterhole buffer
w_dist10.l <- spDists(lion10, water_buffer) # 1:10
w_dist15.l <- spDists(lion15, water_buffer) # 1:15
w_dist20.l <- spDists(lion20, water_buffer) # 1:20

# Creatng a new distance to waterhole column, applying distances generated in matrix
lion10$dist_water <- apply(w_dist10.l, 1, min) # 1:10
lion15$dist_water <- apply(w_dist15.l, 1, min) # 1:15
lion20$dist_water <- apply(w_dist20.l, 1, min) # 1:20
```

##### Hyena to waterhole
```{r}
# Creating a matrix of distances between observation points and the waterhole buffer
w_dist10.h <- spDists(hyena10, water_buffer) # 1:10
w_dist15.h <- spDists(hyena15, water_buffer) # 1:15
w_dist20.h <- spDists(hyena20, water_buffer) # 1:20

# Creatng a new distance to waterhole column, applying distances generated in matrix
hyena10$dist_water <- apply(w_dist10.h, 1, min) # 1:10
hyena15$dist_water <- apply(w_dist15.h, 1, min) # 1:15
hyena20$dist_water <- apply(w_dist20.h, 1, min) # 1:20
```

***
#### Extracting the distance to the nearest road

##### Creating a buffer
```{r}
# Creating a 10 metre buffer around roads
road_buffer <- gBuffer(roads, byid=TRUE, width=10)
# Setting CRS
proj4string(road_buffer) <- projection
```

##### Lion to road
```{r}
# Creating a matrix of distances between observation points and the road buffer
r_dist10.l <- spDists(lion10, road_buffer) # 1:10
r_dist15.l <- spDists(lion15, road_buffer) # 1:15
r_dist20.l <- spDists(lion20, road_buffer) # 1:20

# Creatng a new distance to road column, applying distances generated in matrix
lion10$dist_road <- apply(r_dist10.l, 1, min) # 1:10
lion15$dist_road <- apply(r_dist15.l, 1, min) # 1:15
lion20$dist_road <- apply(r_dist20.l, 1, min) # 1:20
```

##### Hyena to road
```{r}
# Creating a matrix of distances between observation points and the road buffer
r_dist10.h <- spDists(hyena10, road_buffer) # 1:10
r_dist15.h <- spDists(hyena15, road_buffer) # 1:15
r_dist20.h <- spDists(hyena20, road_buffer) # 1:20

# Creatng a new distance to road column, applying distances generated in matrix
hyena10$dist_road <- apply(r_dist10.h, 1, min) # 1:10
hyena15$dist_road <- apply(r_dist15.h, 1, min) # 1:15
hyena20$dist_road <- apply(r_dist20.h, 1, min) # 1:20
```

***
#### Extracting habitat type at each point

```{r}
# Fixing CRS
proj4string(habitat) <- projection
```

##### Lions
```{r}
# Extracting cover type at each point
l10_hab <- over(lion10, habitat) # 1:10
l15_hab <- over(lion15, habitat) # 1:15
l20_hab <- over(lion20, habitat) # 1:20
```

```{r}
# Creating a new column called 'hab_type' with the character column from the extraction
lion10$hab_type <- l10_hab[,2]
lion15$hab_type <- l15_hab[,2]
lion20$hab_type <- l20_hab[,2]
```

##### Hyenas
```{r}
# Extracting cover type at each point
h10_hab <- over(hyena10, habitat) # 1:10
h15_hab <- over(hyena15, habitat) # 1:15
h20_hab <- over(hyena20, habitat) # 1:20
```

```{r}
# Creating a new column called 'hab_type' with the character column from the extraction
hyena10$hab_type <- h10_hab[,2]
hyena15$hab_type <- h15_hab[,2]
hyena20$hab_type <- h20_hab[,2]
```

***
#### Calculating distance to the nearest competitor

##### Converting shapefile data into data frame
```{r}
# Lions
lion10 <- as.data.frame(lion10)
lion15 <- as.data.frame(lion15)
lion20 <- as.data.frame(lion20)

# Hyenas
hyena10 <- as.data.frame(hyena10)
hyena15 <- as.data.frame(hyena15)
hyena20 <- as.data.frame(hyena20)
```

##### Creating a new column in lion and hyena data
```{r}
# Adding new column called 'dist_comp'
# Lions
lion10$dist_comp <- NA
lion15$dist_comp <- NA
lion20$dist_comp <- NA

# Hyenas
hyena10$dist_comp <- NA
hyena15$dist_comp <- NA
hyena20$dist_comp <- NA
```

***
##### Lions 1:10
```{r}
# Creating a date object
dates <- unique(lion10$timestamp)
for(i in 1:length(dates)){

  # subset hyena within 12 hours of lion timestamps
  hyena.temp <- hyena10[hyena10$timestamp > (dates[i] - 60*60*12) & 
                          hyena10$timestamp < (dates[i] + 60*60*12) & hyena10$Presence == 1 ,]
  
  if(nrow(hyena.temp) > 0){
      
      # calculating minimum distance (meters) between lion and hyena - use Pythagoras
      lion10$dist_comp[i] <- min(sqrt((lion10$Easting[i] - hyena.temp$Easting)^2 + (lion10$Northing[i] - hyena.temp$Northing)^2 ), na.rm=T) 

  } else{
    
    lion10$dist_comp[i] = NA
    }# end if loop
} # end i loop
```

##### Lions 1:15
```{r}
# Creating a date object
dates <- unique(lion15$timestamp)
for(i in 1:length(dates)){

  # subset hyena within 12 hours of lion timestamps
  hyena.temp <- hyena15[hyena15$timestamp > (dates[i] - 60*60*12) & 
                          hyena15$timestamp < (dates[i] + 60*60*12) & hyena15$Presence == 1 ,]
  
  if(nrow(hyena.temp) > 0){
      
      # calculating minimum distance (meters) between lion and hyena - use Pythagoras
      lion15$dist_comp[i] <- min(sqrt((lion15$Easting[i] - hyena.temp$Easting)^2 + (lion15$Northing[i] - hyena.temp$Northing)^2 ), na.rm=T) 

  } else{
    
    lion15$dist_comp[i] = NA
    }# end if loop
} # end i loop
```

##### Lions 1:20
```{r}
# Creating a date object
dates <- unique(lion20$timestamp)
for(i in 1:length(dates)){

  # subset hyena within 12 hours of lion timestamps
  hyena.temp <- hyena20[hyena20$timestamp > (dates[i] - 60*60*12) & 
                          hyena20$timestamp < (dates[i] + 60*60*12) & hyena20$Presence == 1 ,]
  
  if(nrow(hyena.temp) > 0){
      
      # calculating minimum distance (meters) between lion and hyena - use Pythagoras
      lion20$dist_comp[i] <- min(sqrt((lion20$Easting[i] - hyena.temp$Easting)^2 + (lion20$Northing[i] - hyena.temp$Northing)^2 ), na.rm=T) 

  } else{
    
    lion20$dist_comp[i] = NA
    }# end if loop
} # end i loop
```

***
##### Hyenas 1:10
```{r}
# Creating a date object
dates <- unique(hyena10$timestamp)
for(i in 1:length(dates)){

  # subset hyena within 12 hours of lion timestamps
  lion.temp <- lion10[lion10$timestamp > (dates[i] - 60*60*12) & 
                          lion10$timestamp < (dates[i] + 60*60*12) & lion10$Presence == 1 ,]
  
  if(nrow(lion.temp) > 0){
      
      # calculating minimum distance (meters) between lion and hyena - use Pythagoras
      hyena10$dist_comp[i] <- min(sqrt((hyena10$Easting[i] - lion.temp$Easting)^2 + (hyena10$Northing[i] - lion.temp$Northing)^2 ), na.rm=T) 

  } else{
    
    hyena10$dist_comp[i] = NA
    }# end if loop
} # end i loop
```

##### Hyenas 1:15
```{r}
# Creating a date object
dates <- unique(hyena15$timestamp)
for(i in 1:length(dates)){

  # subset hyena within 12 hours of lion timestamps
  lion.temp <- lion15[lion15$timestamp > (dates[i] - 60*60*12) & 
                          lion15$timestamp < (dates[i] + 60*60*12) & lion15$Presence == 1 ,]
  
  if(nrow(lion.temp) > 0){
      
      # calculating minimum distance (meters) between lion and hyena - use Pythagoras
      hyena15$dist_comp[i] <- min(sqrt((hyena15$Easting[i] - lion.temp$Easting)^2 + (hyena15$Northing[i] - lion.temp$Northing)^2 ), na.rm=T) 

  } else{
    
    hyena15$dist_comp[i] = NA
    }# end if loop
} # end i loop
```

##### Hyenas 1:20
```{r}
# Creating a date object
dates <- unique(hyena20$timestamp)
for(i in 1:length(dates)){

  # subset hyena within 12 hours of lion timestamps
  lion.temp <- lion20[lion20$timestamp > (dates[i] - 60*60*12) & 
                          lion20$timestamp < (dates[i] + 60*60*12) & lion20$Presence == 1 ,]
  
  if(nrow(lion.temp) > 0){
      
      # calculating minimum distance (meters) between lion and hyena - use Pythagoras
      hyena20$dist_comp[i] <- min(sqrt((hyena20$Easting[i] - lion.temp$Easting)^2 + (hyena20$Northing[i] - lion.temp$Northing)^2 ), na.rm=T) 

  } else{
    
    hyena20$dist_comp[i] = NA
    }# end if loop
} # end i loop
```