---
title: "Lion Modelling"
author: "Sarah Tonnar"
date: "07/07/2023"
output: html_document
---

## Lion Models

### Changing the class of the 'Presence' column
```{r}
lion10$Presence <- as.integer(lion10$Presence)
lion15$Presence <- as.integer(lion15$Presence)
lion20$Presence <- as.integer(lion20$Presence)
```

### Rescaling numeric variables around 0
```{r}

lion10s <- lion10
lion15s <- lion15
lion20s <- lion20

# Lion 1:10
lion10s$dist_water <- scale(lion10$dist_water)
lion10s$dist_road <- scale(lion10$dist_road)
lion10s$dist_comp <- scale(lion10$dist_comp)

# Lion 1:15
lion15s$dist_water <- scale(lion15$dist_water)
lion15s$dist_road <- scale(lion15$dist_road)
lion15s$dist_comp <- scale(lion15$dist_comp)

# Lion 1:20
lion20s$dist_water <- scale(lion20$dist_water)
lion20s$dist_road <- scale(lion20$dist_road)
lion20s$dist_comp <- scale(lion20$dist_comp)
```

### Variables Only

```{r}
# Creating a model with all variables in it and ID as a random effect for 1:15
lion10mod <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + (1|ID),
                    family=binomial, data = na.omit(lion10s))
```

```{r}
# Creating a model with all variables in it and ID as a random effect for 1:15
lion15mod <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + (1|ID),
                    family=binomial, data = na.omit(lion15s))
```

```{r}
# Creating a model with all variables in it and ID as a random effect for 1:15
lion20mod <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + (1|ID),
                    family=binomial, data = na.omit(lion20s))
```

#### Removing dist_water
```{r}
lion10mod1 <- glmer(Presence ~ dist_road + hab_type + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion10s))
```

```{r}
lion15mod1 <- glmer(Presence ~ dist_road + hab_type + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion15s))
```

```{r}
lion20mod1 <- glmer(Presence ~ dist_road + hab_type + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10mod1, lion10mod)
lrtest(lion15mod1, lion15mod)
lrtest(lion20mod1, lion20mod)

# P < 0.05 - we will keep dist_water in the model
```

#### Removing dist_road
```{r}
lion10mod2 <- glmer(Presence ~ dist_water + hab_type + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion10s))
```

```{r}
lion15mod2 <- glmer(Presence ~ dist_water + hab_type + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion15s))
```

```{r}
lion20mod2 <- glmer(Presence ~ dist_water + hab_type + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10mod2, lion10mod)
lrtest(lion15mod2, lion15mod)
lrtest(lion20mod2, lion20mod)

# P < 0.05 - we will keep dist_road in the model
```

#### Removing hab_type
```{r}
lion10mod3 <- glmer(Presence ~ dist_water + dist_road + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion10s))
```

```{r}
lion15mod3 <- glmer(Presence ~ dist_water + dist_road + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion15s))
```

```{r}
lion20mod3 <- glmer(Presence ~ dist_water + dist_road + dist_comp + (1|ID), family=binomial,
                    data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10mod3, lion10mod)
lrtest(lion15mod3, lion15mod)
lrtest(lion20mod3, lion20mod)

# P < 0.05 - we will keep hab_type in the model
```

#### Removing dist_comp
```{r}
lion10mod4 <- glmer(Presence ~ dist_water + dist_road + hab_type + (1|ID), family=binomial,
                    data=na.omit(lion10s))
```

```{r}
lion15mod4 <- glmer(Presence ~ dist_water + dist_road + hab_type + (1|ID), family=binomial,
                    data=na.omit(lion15s))
```

```{r}
lion20mod4 <- glmer(Presence ~ dist_water + dist_road + hab_type + (1|ID), family=binomial,
                    data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10mod4, lion10mod)
lrtest(lion15mod4, lion15mod)
lrtest(lion20mod4, lion20mod)

# P < 0.05 - we will keep dist_comp in the model
```

## Adding interactions

#### Adding Season:dist_water
```{r}
lion10season <- glmer(Presence ~ dist_water + dist_road + hab_type + Season:dist_water +
                        dist_comp + (1|ID), family=binomial, data=na.omit(lion10s))
```

```{r}
lion15season <- glmer(Presence ~ dist_water + dist_road + hab_type + Season:dist_water +
                        dist_comp + (1|ID), family=binomial, data=na.omit(lion15s))
```

```{r}
lion20season <- glmer(Presence ~ dist_water + dist_road + hab_type + Season:dist_water +
                       dist_comp +  (1|ID), family=binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10mod, lion10season)
lrtest(lion15mod, lion15season)
lrtest(lion20mod, lion20season)

# P < 0.05 - keep Season:dist_water in the model
```

#### Adding Season:dist_road
```{r}
lion10season1 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + (1|ID), family=binomial, data=na.omit(lion10s))
```

```{r}
lion15season1 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + (1|ID), family=binomial, data=na.omit(lion15s))
```

```{r}
lion20season1 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + (1|ID), family=binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season, lion10season1)
lrtest(lion15season, lion15season1)
lrtest(lion20season, lion20season1)

# P < 0.05 - keep in model
```

#### Adding Season:hab_type
```{r}
lion10season2 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion10s))
```

```{r}
lion15season2 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion15s))
```

```{r}
lion20season2 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season1, lion10season2)
lrtest(lion15season1, lion15season2)
lrtest(lion20season1, lion20season2)

# P < 0.05 - keep Season:hab_type in the model
```

#### Adding Season:dist_comp
```{r}
lion10season3 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + (1|ID),
                       family = binomial, data=na.omit(lion10s))
```

```{r}
lion15season3 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + (1|ID),
                       family = binomial, data=na.omit(lion15s))
```

```{r}
lion20season3 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + (1|ID),
                       family = binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season2, lion10season3)
lrtest(lion15season2, lion15season3)
lrtest(lion20season2, lion20season3)

# P < 0.05 - keep in model
```

#### Adding DayNight:dist_water
```{r}
lion10season4 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_water +
                         (1|ID), family = binomial, data=na.omit(lion10s))
```

```{r}
lion15season4 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_water +
                         (1|ID), family = binomial, data=na.omit(lion15s))
```

```{r}
lion20season4 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_water +
                         (1|ID), family = binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season3, lion10season4)
lrtest(lion15season3, lion15season4)
lrtest(lion15season3, lion15season4)

# P > 0.05 - remove from model
```

#### Adding DayNight:dist_road
```{r}
lion10season5 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         (1|ID), family = binomial, data=na.omit(lion10s))
```

```{r}
lion15season5 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         (1|ID), family = binomial, data=na.omit(lion15s))
```

```{r}
lion20season5 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         (1|ID), family = binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season3, lion10season5)
lrtest(lion15season3, lion15season5)
lrtest(lion20season3, lion20season5)

# P < 0.05 - keep in model
```

#### Adding DayNight:hab_type
```{r}
lion10season6 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         DayNight:hab_type + (1|ID), family = binomial, data=na.omit(lion10s))
```

```{r}
lion15season6 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         DayNight:hab_type + (1|ID), family = binomial, data=na.omit(lion15s))
```

```{r}
lion20season6 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         DayNight:hab_type + (1|ID), family = binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season5, lion10season6)
lrtest(lion15season5, lion15season6)
lrtest(lion20season5, lion20season6)

# P < 0.05 - keep in model
```

#### Adding DayNight:dist_comp
```{r}
lion10season7 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         DayNight:dist_comp + DayNight:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion10s))
```

```{r}
lion15season7 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         DayNight:dist_comp + DayNight:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion15s))
```

```{r}
lion20season7 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         DayNight:dist_comp + DayNight:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season6, lion10season7)
lrtest(lion15season6, lion15season7)
lrtest(lion20season6, lion20season7)

# P > 0.05 - remove from model
```

#### Adding dist_water:dist_road
```{r}
lion10season8 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion10s))
```

```{r}
lion15season8 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion15s))
```

```{r}
lion20season8 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + (1|ID), family = binomial,
                       data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season6, lion10season8)
lrtest(lion15season6, lion15season8)
lrtest(lion20season6, lion20season8)

# P < 0.05 - keep in model
```

#### Adding dist_water:hab_type
```{r}
lion10season9 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + (1|ID), 
                       family = binomial, data=na.omit(lion10s))
```

```{r}
lion15season9 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + (1|ID), 
                       family = binomial, data=na.omit(lion15s))
```

```{r}
lion20season9 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + (1|ID), 
                       family = binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season8, lion10season9)
lrtest(lion15season8, lion15season9)
lrtest(lion20season8, lion20season9)

# P < 0.05 - keep in model
```

#### Adding dist_water:dist_comp
```{r}
lion10season10 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + (1|ID),family = binomial, data=na.omit(lion10s))
```

```{r}
lion15season10 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + (1|ID),family = binomial, data=na.omit(lion15s))
```

```{r}
lion20season10 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + (1|ID),family = binomial, data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season9, lion10season10)
lrtest(lion15season9, lion15season10)
lrtest(lion20season9, lion20season10)

# P < 0.05 - keep in model
```

#### Adding dist_road:hab_type
```{r}
lion10season11 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + (1|ID), family = binomial,
                        data=na.omit(lion10s))
```

```{r}
lion15season11 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + (1|ID), family = binomial,
                        data=na.omit(lion15s))
```

```{r}
lion20season11 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + (1|ID), family = binomial,
                        data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season10, lion10season11)
lrtest(lion15season10, lion15season11)
lrtest(lion20season10, lion20season11)

# P < 0.05 - keep in model
```

#### Adding dist_road:dist_comp
```{r}
lion10season12 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + dist_road:dist_comp + (1|ID), 
                        family = binomial,data=na.omit(lion10s))
```

```{r}
lion15season12 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + dist_road:dist_comp + (1|ID), 
                        family = binomial,data=na.omit(lion15s))
```

```{r}
lion20season12 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + dist_road:dist_comp + (1|ID), 
                        family = binomial,data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season11, lion10season12)
lrtest(lion15season11, lion15season12)
lrtest(lion20season11, lion20season12)

# P > 0.05 - remove from model
```

#### Adding hab_type:dist_comp
```{r}
lion10season13 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + hab_type:dist_comp + (1|ID), 
                        family = binomial,data=na.omit(lion10s))
```

```{r}
lion15season13 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + hab_type:dist_comp + (1|ID), 
                        family = binomial,data=na.omit(lion15s))
```

```{r}
lion20season13 <- glmer(Presence ~ dist_water + dist_road + hab_type + dist_comp + Season:dist_water +
                         Season:dist_road + Season:hab_type + Season:dist_comp + DayNight:dist_road +
                         dist_water:dist_road + DayNight:hab_type + dist_water:hab_type + 
                         dist_water:dist_comp + dist_road:hab_type + hab_type:dist_comp + (1|ID), 
                        family = binomial,data=na.omit(lion20s))
```

```{r}
# Doing an LRT test
lrtest(lion10season11, lion10season13)
lrtest(lion15season11, lion15season13)
lrtest(lion20season11, lion20season13)

# P > 0.05 - remove from the model
```

## Setting final model
```{r}
# Using 1:15 random point ratio as results were consistent with this number
lion.mod.final <- lion15season11
```
